Bird beak's elaboration and innovation

### 2022-10-12 Analysis/03-Elaboration_innovation_analysis.Rmd
tests for variance (to be removed)

## Get the variance for each level
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = TRUE)
## Rename the columns
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")
reorder_clades <- c(9:35,1:8, 36, 37)
## Pull by term
variance_per_levels<- cbind("orders" = apply(variance_per_clades[, 9:35], 1, sum),
                            "super-orders" = apply(variance_per_clades[, 1:8], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
variance_per_terms <- cbind("clades" = apply(variance_per_clades[, 1:35], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
## Plot the results
pdf(file = "../Manuscript/Figures/variance_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
boxplot(variance_per_terms, horizontal = TRUE, las = 1, main = "Scaled variance by terms", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(variance_per_levels, horizontal = TRUE, las = 1, main = "Scaled variance by levels", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(variance_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled variance by clades", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
dev.off()

>>>>

## Get the variance for each level
variance_per_clades <- MCMCglmm.variance(shapespace_allbirds_lvl_superorder_order, scale = TRUE)
## Rename the columns
colnames(variance_per_clades) <- c(levels(shapespace$level1)[-1], levels(shapespace$level2)[-1], "phylogeny", "residuals")
reorder_clades <- c(9:35,1:8, 36, 37)
## Pull by term
variance_per_levels<- cbind("orders" = apply(variance_per_clades[, 9:35], 1, sum),
                            "super-orders" = apply(variance_per_clades[, 1:8], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
variance_per_terms <- cbind("clades" = apply(variance_per_clades[, 1:35], 1, sum),
                            "phylogeny" = variance_per_clades[, 36],
                            "residuals" = variance_per_clades[, 37])
## Plot the results
pdf(file = "../Manuscript/Figures/variance_per_levels_whole_tree.pdf", height = 8, width = 14)
par(mar = c(5,8,2,2), mfrow = c(1, 3))
# boxplot(variance_per_terms, horizontal = TRUE, las = 1, main = "Scaled variance by terms", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
# boxplot(variance_per_levels, horizontal = TRUE, las = 1, main = "Scaled variance by levels", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(variance_per_clades[,reorder_clades], horizontal = TRUE, las = 1, main = "Scaled variance by clades", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(log(variance_per_clades[,reorder_clades])/log(sample_sizes), horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing sample size effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
boxplot(log(variance_per_clades[,reorder_clades])/log(observed_variance)[names(variances)], horizontal = TRUE, las = 1, main = "Scaled variance by clades (removing variance effect)", xlab = "Scaled variance\n(sum of each VCV variance part (diagonal))")
dev.off()

## Getting the variances per clades
variances <- apply(variance_per_clades[,reorder_clades], 2, mean)

## Getting the sample sizes per clades (the one for the residuals is the whole tree size)
sample_sizes <- c("residuals" = size.subsets(order_shapespace)[28],
                  size.subsets(superorder_shapespace),
                  size.subsets(order_shapespace))[names(variances)]

## Measuring the variance for each clade and for the residuals
## the obs variance of the residuals is just the variance in the whole dataset (phylogeny)
observed_variance <- c("residuals" = unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[28], 
                       unlist(get.disparity(dispRity(superorder_shapespace, metric = sd))),
                       unlist(get.disparity(dispRity(order_shapespace, metric = sd)))[-28])^2

plot(log(variances)/log(sample_sizes), log(sample_sizes), xlab = "log variance / log n", ylab = "log n")

x <- rnorm(100)
y <- rnorm(100)
z <- rnorm(100)

cor(x, y)
cor(x/y, y)
